---
title: "Formative Assessment 3"
output: rmarkdown::github_document
---

#### Probability and Probability Distribution

#### By Romand Lansangan

### 1)
A binary communication channel carries data as one of two sets of signals denoted by 0 and 1. Owing to noise, 
a transmitted 0 is sometimes received as a 1, and a transmitted 1 is sometimes received as a 0. For a given channel, 
it can be assumed that a transmitted 0 is correctly received with probability 0.95, and a transmitted 1 is correctly 
received with probability 0.75. Also, 70% of all messages are transmitted as a 0. If a signal is sent, 
determine the probability that:
```
(a) a 1 was received;
(b) a 1 was transmitted given than a 1 was received.
```

#### Data:
``` {r}
dataTable <- data.frame(
  message = c(0, 1),
  trans_prob = c(0.70, 0.30),
  rec_prob = c(0.95, 0.75)
)
dataTable
```

Where trans_prob is the probability of a message being transmitted and rec_prob as the probability of them being received correctly.

#### Tree:
To illustrate the problem in a more feasible manner, a tree diagram were created.
``` {r}
library(igraph)

vertices <- c("Sender", "0", "1", "0", "1", "0", "1", "Receiver")

edges <- matrix(c(
  "Sender", "0",
  "Sender", "1",
  "0", "0_0",
  "0", "0_1",
  "1", "1_0",
  "1", "1_1",
  "0_0", "Receiver",
  "0_1", "Receiver",
  "1_0", "Receiver",
  "1_1", "Receiver"
), byrow = TRUE, ncol = 2)

edge_labels <- c("0.70", "0.30", "0.95", "0.05", "0.25","0.75", "", "", "", "")

g <- graph_from_edgelist(edges, directed = TRUE)
E(g)$label <- edge_labels

V(g)$label <- vertices

layout <- matrix(c(
  -0.5, 0, #sender
  -2, -1, #0
  1, -1, #1
  -3, -2,#0_0
  -1, -2, #0_1
  0, -2,#1_0
  2, -2, #1_1
  -0.5, -3 #Receiver
), ncol = 2, byrow = TRUE)

V(g)$x <- layout[,1]
V(g)$y <- layout[,2]

par(mar = c(0, 0, 3, 0))

plot(g, layout = layout, edge.label = E(g)$label, edge.label.cex = 0.8, vertex.label.cex = 0.8,
     vertex.shape = "circle", vertex.size = 30, edge.arrow.size = 0.5, 
     main = "Tree Diagram of Probability for Binary Communication Channel")

label_coords <- matrix(c(
  -0.95, 0.66,
  1.4, -0.01
), ncol = 2, byrow = TRUE)



text(label_coords[1,1], label_coords[1,2], labels = "Probability of being TRANSMITTED", pos = 3, cex = 0.8)
text(label_coords[2,1], label_coords[2,2], labels = "Probability of being RECEIVED", pos = 3, cex = 0.8)

```

In this tree diagram, the root is the "sender" wherein the branching edges (and their respective label) represents the probability that it will be sent. In this case, the "sender" has a 0.70 (or 70%) chance of transmitting "0", thus leaving the rest of possible probability (0.30 or 30%) to transmitting "1".

The nodes at the end of each edges represents the message that was sent. The edges that branched out from them represents the probability of a message being received. In this case, when the "sender" sent "0," there is a 0.05 (5%) chance that it will be received by the receiver as "1"—and so on.

#### A) 1 was received.

To calculate the probability of message being correctly (and wrongly) received, we shall use the Multiplication Law of Probability. 


##### Multiplication Law of Probability:

The Multiplication Law of Probability is defined as the following:
$$
P(E_{1} \cap E_{2}) = P(E_{2}|E_{1})P(E_{1})
$$

Let $E_{1}$ be an event in which "1" was received;
Let $A$ be the case in which the "sender" sent 0;
Let $B$ be the case in which the "sender" sent 1.

For calculating the probability of "1" being received, we will have 2 possibilities. One being that the "sender" sent "1" and it was correctly received ($P(B \cap E_{1})$) and the "sender" transmitting "0" and was incorrectly received ($P(A \cap E_{1})$).
With all that in mind we'll have the following:

The probability of event $A$ intersecting with event $E_{1}$ is given by:

$$
P(A \cap E_{1}) = P(E_{1}|A)P(A)
$$

##### Solution:

Substituting the values we get:

$$
P(A \cap E_{1}) = 0.05 \times 0.70
$$

$$
P(A \cap E_{1}) = 0.035
$$

Similarly, for events $B$ and $E$:

$$
P(B \cap E_{1}) = P(E_{1}|B)P(B)
$$

And with values:

$$
P(B \cap E_{1}) = 0.75 \times 0.30
$$

$$
P(B \cap E_{1}) = 0.225
$$

```{r}
edge_labels <- c("0.70", "0.30", "0.95", "0.05", "0.25","0.75", "", "0.035", "", "0.225")

E(g)$label <- edge_labels

plot(g, layout = layout, edge.label = E(g)$label, edge.label.cex = 0.8, vertex.label.cex = 0.8,
     vertex.shape = "circle", vertex.size = 30, edge.arrow.size = 0.5, 
     main = "Tree Diagram of Probability for Binary Communication Channel")

label_coords <- matrix(c(
  -1, 0.48,
  1.5, -0.18
), ncol = 2, byrow = TRUE)



text(label_coords[1,1], label_coords[1,2], labels = "Probability of being TRANSMITTED", pos = 3, cex = 0.8)
text(label_coords[2,1], label_coords[2,2], labels = "Probability of being RECEIVED", pos = 3, cex = 0.8)

```

Since we are calculating for the probability that a "1" was received, the "receiver" must only receive one and it doesn't particularly matter if it either "0" or "1" was originally transmitted. With this in mind, because the two cases mentioned above were mutually exclusive, we can use the Law of Total Probability.

##### Law of Total Probability:

The Law of Total Probability is defined as the following:  

If a sample space $S$ can be partitioned into k mutually exclusive and exhaustive events,
$A_{1}$, $A_{2}$, $A_{3}$, … , $A_{k}$, then for any event E:

$$
P(E) = P(A_{1})P(E|A_{1}) + P(A_{2})P(E|A_{2}) + P(A_{3})P(E|{A_{3}}) + ... P(A_{k})P(E|A_{k})
$$


With that, we'll have
$$
P(E_{1}) = P(E_{1}|A)P(A) + P(E_{1}|B)P(B)
$$

$$
P(E_{1}) = 0.035 + 0.225
$$

Thus, the total probability of receiving a "1" is:
$$
P(E_{1}) = 0.26
$$

#### B) 1 was transmitted given than a 1 was received.

Let $E_{1}$ be an event in which "1" was received;
Let $B$ be the case in which the "sender" transmitted 1.

We can rewrite this particular problem as a conditional probability:
$$
P(B|E_{1}) = 0.26
$$

##### Conditional Probability:
The conditional probability is defined as
$$
P(E_{2}|E_{1}) = \frac{P(E_{2} \cap E_{1})}{P(E_{1})}
$$

##### Solution:

Applying the formula to the problem, 

$$
P(B|E_{1}) = \frac{P(B \cap E_{1})}{P(E_{1})}
$$

Substituting the values we get:
$$
P(B|E_{1}) = \frac{0.225}{0.26}
$$
Thus, the probability that "1" was transmitted given than a "1" was received is
$$
P(B|E_{1}) \approx 0.8653
$$